\documentclass{article}

\usepackage{hyperref}

\title{PanzerBlitz Rules of Play}

\begin{document}

\maketitle

\tableofcontents

General Outline of Play ................................................2
The Mapboard ...........................................................2
The Playing Pieces .....................................................2
Factor Definitions .....................................................2
Unit Identification Table and the Program Identity Code System (PICS) ..2


Movement................................................................3 
Terrain Features .......................................................3 
How to Move Units.......................................................3 
Transporting Units ..................................................................................................................................................3 
Road Movement .....................................................................................................................................................4 
Stacking (more than one unit per hex) ...................................................................................................................4 


Combat .......................................................................................................................................................................5 
How To Have Combat ............................................................................................................................................5 
Overrun Attack ........................................................................................................................................................5 
Close Assault Tactics .............................................................................................................................................6 
Weapon-to-Target Relationships ............................................................................................................................6 
WEC Weapons Effectiveness Chart .......................................................................................................................7 


Obstacles and Elevations...........................................................................................................................................7 
How to Determine the Line of Sight/ Line of Fire (LOS/LOF) .................................................................................7 
Special Notes (To Be Used In Conjunction With TET) ...........................................................................................8 
Hill and Slope Defense Exceptions ........................................................................................................................8 
Gullies and Streambeds .........................................................................................................................................9 
Spotting...................................................................................................................................................................9 
Wreckage................................................................................................................................................................9 
Mines ......................................................................................................................................................................9 


Positional Defenses..................................................................................................................................................10 
Blockage ...............................................................................................................................................................10 
Fortifications .........................................................................................................................................................10 


Game Procedure ......................................................................................................................................................11 
Sequence of Play..................................................................................................................................................11 
German Player Turn .............................................................................................................................................11 
Russian Player Turn .............................................................................................................................................11 


Optional Rules ..........................................................................................................................................................11 
Indirect Fire ...........................................................................................................................................................11 
Real Space Line of Sight Determination ..............................................................................................................11 
Panzerblitz Assault ...............................................................................................................................................11 
Ammunition Rule ..................................................................................................................................................12 


Experimental Rules ..................................................................................................................................................12 
Hidden Deployment ..............................................................................................................................................12 
Experimental Indirect Fire.....................................................................................................................................12 



\section{General Outline of Play}

Each side maneuvers its forces (playing pieces) on
the terrain map seeking to destroy the enemy units
and/or gain a specific territorial objective as outlined
in one of the scene-setting Situation Cards. Players
move their pieces and have combat by taking turns.
Each complete turn represents six minutes of real
time.

The forces in a given Situation may be unequal and
one side may have a better chance of winning than
the other, but it is primarily the skill of the individual
player which determines the outcome of the game.
The chance element introduced by the use of the
die-roll/Combat Results Table is only that degree of
chance consistently present in any real-life combat
event. The probabilities of combat outcomes have
been worked out using historical and technical data.

\section{The Mapboard}

The three section mapboard represents a varied
sampling of typical terrain in the Soviet Union. Each
section has a number (1, 2 or 3) located just above
the fold-line. The board is ``geo-morphic'' i.e. capable
of being changed by re-arranging the three sections
in a variety of juxtapositions. The long edge of each
two panel section will line up with the long edge of
any other section no matter which way they are
butted together. The short edges mate with each
other in a similar fashion.

\section{The Playing Pieces}

The square, cardboard pieces represent platoon or
company sized military units of several different
types (e.g. Infantry platoons, Tank platoons, Assault
Gun platoons, etc), which are the playing pieces
used in PANZERBLITZ. Hereafter they will be
referred to as “units” or “unit counters”. The numbers
on the unit-counters represent that unit’s capabilities
with respect to movement, attack, defense, and
range of weapons. The other symbols or silhouettes
identify what type of unit that counter represents.


Notice that all vehicle units are symbolized with an 
appropriate silhouette and all other (non-vehicular) 
units are symbolized with standard military-planning 
symbols.

\section{Factor Definitions}

MOVEMENT FACTOR (MF) – The basic, maximum
number of hexagons (hexes) which a unit may move in
one turn. This capability can be reduced or increased
by terrain features.

ATTACK FACTOR (AF) – The basic offensive power of a
given unit.

DEFENSE FACTOR (DF) – The basic defensive power of 
a given unit.

RANGE FACTOR (RF) – The maximum effective 
distance (in hexagons) that a unit’s Attack Factor can 
be used against enemy units. 

For example, a unit with a RF of 8 could fire its 
weapons (use its attack factor) against any enemy unit 
within that 8 hex range. 

For a full display of all the counters in PanzerBlitz see 
the Unit Identification Table. 

Unit Identification Table and the Program 
Identity Code System (PICS) 

The Unit Identification Table shows a full “breakout” of 
all the units in PanzerBlitz grouped according to 
general category and function. Unless otherwise 
stated, all Russian units are COMPANIES and all 
German units are PLATOONS. The Unit Composition 
portion of the table shows what actually went into the 
make-up of the various units. (Note: Although the non0vehicular 
units had organic transport assigned to 
them, the game-counters symbolize them without this 
transport – the trucks and wagons being given as 
separate counters in each Situation.) 

Each specific unit type has been assigned a specific 
code number (example: Hetzer No. 832). Each 



specific TYPE of unit has a common second digit 
code number (example: all German “Hetzers” have 
the number “3” as their second digit). Each 
FUNCTIONAL CATEGORY is expressed by the left-
most number (example: all Tank Destroyers are in 
the 800 series). 

Roughly comparable Russian and German units 
have been assigned the same TYPE code number 
group. Towed guns have been assigned two-digit 
numbers (although they may be thought of as having 
a FUNCTIONAL CODE of “0”). To fully identify a 
given unit, write a “G” or “R” (nationally) followed by 
its PIC number: Thus “G/832” means: German/TankDestroyer/
Hetzer/2nd Platoon. 

Note: the right hand digit will only be a zero if there 
is only one counter of that type supplied. The 
second digit will only be a zero if there is only one 
TYPE in that category. 

The Program Identity Code system is simply a 
shorthand method of positive unit identification for 
use in play-by-mail games and noting the position of 
units in games which must be interrupted and restarted. 


\section{Movement}

Terrain Features 

The hexagonal grid superimposed upon the 
mapboard is used to determine movement and to 
delineate the boundaries of the various terrain 
features. A hex is considered to be a given type of 
terrain if all or any part of it contains that terrain 
feature. Terrain affects movement and defense as 
outlined in the Terrain Effects Chart (TEC). 

The “heavy-hex-side” symbols (different colored 
bars superimposed upon some hex-sides) are 
explained in the OBSTACLE AND ELEVATIONS 
section of the rules. 

The single hexagon on board section No. 3 which is 
completely covered by the pond is unenterable for 
ALL units. Units may use the hexes partially covered 
by the pond. 

The half-hexes on the outer edges of the board are 
considered playable and may be utilized as if they 
were complete hexagons. 

How to Move Units 

A- 
In any one turn a player may move as many or as 
few of his units as he desires. 

B- 
Units which have fired (used their AF) may not 
move in that turn. “Dispersed” units may not move 
(see How to Have Combat and the Combat 
Results Table). 

C-Units may move as much or as little as the player 
desires within the limits of their MF’s and the 
terrain effects. Units with a MF of 1 may move one 
hex per turn regardless of terrain. 

D- 
Units may move through friendly units. 

E- 
Units may not move through enemy units 
(Exception: see OVERRUN RULE) 

F- 
Units may not stop on top of enemy units. 

G- 
There is no movement penalty when moving into 
or through hexes adjacent to enemy units (i.e. 
there is no “zone of control” such as in other 
games. 

H- 
No enemy movement is allowed during friendly 
movement. 

I- 
No combat, enemy or friendly, takes place during 
movement (Exception: see OVERRUN RULE) 

Transporting Units 

A- 
The “C” class vehicular units (wagons, trucks and 
halftracks have the capability of carrying non-
vehicular units (guns, infantry and command 
posts). Each “C” unit has the capacity to carry one 
non-vehicular unit. To symbolize that a unit is a 
passenger in a “C” unit, place the unit being 
carried UNDER the “C” unit. Players should never 
place non-vehicular units under vehicular units 
unless they are being transported by that unit. 

B-In any one turn a “C” unit may either “Load” or 
“Transport” or “Transport and Unload.” It may only 
perform one of these operations per turn. 

C- 
“C” units and the unit(s) to be loaded must begin 
their turn on the same hex. The passenger unit 
may not fire (use its AF in the turn of loading, while 
being transported or when unloading. Dispersed 
units may not load or unload. Units are “loaded” 
when they are under the “C” unit. 

D- 
Passengers may not move independently on the 
turn in which they unload from transporting units. 



E- 
A truck or wagon unit an the unit is transporting 
have a combined defense factor of 1. 

F- 
When using halftracks as transport, the defense 
factor (DF) of the halftrack unit is used when 
attacked. Elimination affects both carrier and 
passenger. The halftrack unit may fire while 
loaded. 

G- 
Passengers and carriers are treated as one unit 
for stacking purposes (see Stack Limitations). 
Units are “Loaded” when under the “C” unit. 
Passengers and their carriers are treated as one 
unit for combat results purposes. If combat 
results call for elimination, both are eliminated. If 
dispersed, both are dispersed. 

H- 
Armored vehicle units (tanks, assault guns, etc.) 
may carry non-vehicular units in a fashion 
similar to “C” units. Each armored unit may carry 
one passenger unit. 

I- 
If an armored unit is destroyed while 
transporting, both passenger and carrier are 
destroyed. Units traveling on armored units may 
be attacked exclusive of the armored unit in 
which case the passenger unit has a DF of 1. 
The armored unit is unaffected if only its 
passengers are attacked. 

J- 
Armored units may fire when loading, unloading 
or carrying passengers. All other transport rules, 
however, apply. 

K- 
Russian cavalry units may NOT be transported 
by “C” units or armored units. 

Road Movement 

Units traveling along roads do so at the road 
movement rate regardless of the other terrain in the 
road hexes. All units may travel over all roads of the 
board regardless of accompanying terrain in which 
the might otherwise be prohibited. Roads do not 
alter the defense effects of surrounding terrain. 

A- 
All units move along roads at a cost of . 
movement factor per road hex. Entering a road 
hex through a non-road hex side is done at the 
MF cost of the other terrain in the road hex 
being entered. 

B- 
Units may freely combine road and non-road 
movement in the same turn. 

C- 
Units may not stack while moving along road at 
the road movement rate. 

D- 
Units may not move through or onto other units on 
a road when moving at the road movement rate. 

E- 
To move through or onto a friendly unit on a road 
costs the full non-road movement cost of the other 
terrain in that hex, the moving unit incurs the non-
road MF cost of the other terrain in that hex as 
well. In effect, you are “passing” the unit sitting on 
the road b swinging off the road and maneuvering 
around it. Terrain hexes or hex sides through 
which a unit would be prohibited to travel when off 
the road cannot be traveled upon when performing 
this “passing” maneuver. For example a vehicular 
unit could NOT move through a unit on a 
swamp/road hex. It could, however. move ONTO 
such a unit and move off in the next turn. 

F-Units may stack with other units and ove along 
roads at the NON-road movement rate (and a unit 
in such a stack could split off and move ahead by 
itself at the road movement rate). 

G- 
Remember: a vehicle with passengers is 
considered as one unit and may therefore travel 
together at the road movement rate. 

Stacking (more than one unit per hex) 

A- 
The Russians may stack two units per hex. 

B- 
The Germans may stack three units per hex. 

C- 
When a unit is being carried by another unit, the 
passenger and the carrier is considered as one 
unit for stacking purposes. 

D- 
Stacking limitations do not apply during 
movement. They only apply before and after 
movement (except as qualified by the road 
movement rules). 

E- 
Minefield counters and Fortification counters do 
not count towards stacking limits. 

F- 
Block counters and Wreck counters ARE counted 
towards stacking limits. 

G- Vehicle units may not stack with other vehicle 
units on swamp-road hexes. 

Page 4 of 12 


Combat 

How To Have Combat 

A- 
Basically, to have combat, the attacking unit 
compares its Attack Factor (AF) to the defending 
unit’s Defense Factor (DF). The comparison is 
stated as a ratio: AF to DF; then rounded off in 
the defender’s favor to conform to the ratios 
given on the Combat Results Table (CRT). 
Example: 11 to 3 rounds off to 3 to 1. Roll the 
die and take the action indicated by the CRT. 

B- 
Attack takes place before the movement portion 
of a player’s turn. Only the player whose turn it 
is may attack, the other player is considered the 
“defender.” 

C- 
Only enemy units within the Range Factor (RF) 
of the attacking unit my be fired upon by that 
unit. 

D- 
A player may make as many or as few attacks 
per turn as he desires (within the restrictions of 
the rules of combat). A player is never forced to 
attack. Attacking is an act of volition. 

E- 
Every firing unit firing on the same defending 
unit must combine their Attack Factors into one 
large Attack Factor before computing odds. 
Units may fire only once per turn. 

F- 
Units which fire (attack) in the combat portion of 
the turn may NOT move in the movement 
portion of the same turn. 

G- 
Different Attacking units may fire at the same 
target unit. Each firing unit is announced to be 
firing at a common target, and the combined 
attack is resolved all at once. 

H- 
Units may not split their attack factor (i.e. a given 
attacking unit could not apply part of its factor to 
one attack and part to another). Attack and 
defense factors are not “transferable” from one 
unit to another. Each unit is treated as an 
indivisible set of factors. 

I- 
When there is more than one defending unit in a 
hexagon, attacks against that stack of units may 
be prosecuted in one of the three mutually 
exclusive ways: 

1. SELECTIVE ATTACK: Only one unit (the 
weakest) in the defending stack is attacked; the 
others are ignored and may not be attacked; the 
others are ignored and may not be attacked in the 
normal attack phase of that run. The “weakest” unit 
is defined as being that unit in a defending stack 
against which that particular attacking units(s) can 
attack at the highest odds ratio. Selective attacks 
may be made at odds as low as 1 to 4. 

2. MULTIPLE ATTACK: More than one unit in the 
defending stack is attacked. Each unit attacked is 
treated as a separate battle with the die being 
rolled once for each battle. In this type of attack the 
WEAKEST unit is attacked first at at least 1 to 1 
odds: finally the strongest unit (defense factor) may 
be attacked (at at least 1 to 1). By definition, at 
least TWO of the defending units in a stack must 
be attacked to use this method. Of course, to be 
able to make a Multiple Attack against a defending 
stack, the attacker must use at least one different 
attacking unit against each defender. Remember a 
unit may never use its attack factor more than once 
per turn. 
3. COMBINATION ATTACK: The defending units are 
treated as one combined defense factor and the 
attacker totals his units into one combined attack 
factor. The die is rolled once for the entire stack 
and the result applies to all the defending units in 
that stack. Combination attacks may be made by 
as few as one attacking unit and at odds as low as 
1 to 4. See the Weapons-to-Target Relationships 
rule. 
Overrun Attack 

A- 
Armored vehicle units may overrun enemy units in 
clear terrain. 

B- 
To overrun a unit or stack of units, move the 
attacking armored vehicle unit(s) straight through 
the enemy-occupied hex, exiting into the hex 
DIRECTLY opposite the hex of entry. Overrunning 
units must stop in the “exit-hex” and may move no 
further that turn. If the exit hex is occupied by 
enemy units, the overrun may not be made. 
Overrunning units may not travel at the road-
movement rate during that turn. Overrunning units 
must have sufficient movement factors available to 
reach the exit hex. The exit hex does not have to 
be a clear terrain hex, nor does the entry hex: only 
the target hex must be clear terrain. 

C- 
As you move over the enemy unit or stack of units, 
execute your attack. This is the only case in which 

Page 5 of 12 


an attack may be made during the movement 
phase of a turn. 

D- 
Overrunning units attack with an increased 
combat effect. Figure the odds ratio of the attack 
using the basic AF to DF system then increase 
the odds by one in favor of the attacker ( e.g. a 
ratio of 3 to 1 increases to 4 to 1). Also subtract 
2 from the die roll results (e.g. a die roll of 3 is 
treated as if it were a roll of 1). A defending 
stack is treated as one combined defense factor 
when being overrun. More than one armored 
unit may overrun an enemy stack and the 
overrunning units do not have to enter and exit 
through the same hexes. They must, however, 
combine their attack into one large attack factor. 
In other words, a defending unit or stack may 
not suffer more than one overrun attack per turn. 

E- 
In determining overrun odds use only the factors 
printed on the units (for attacker and defender). 
Do not halve or double the attack factors as 
shown on the Weapons Effectiveness Chart. 
Use only the overrun “bonus” as outlined in rule 
“D”. 

F- 
You may not fire overrunning units during the 
usual combat portion of the turn in which the 
overrun is made. 

G-Units on Block, Wreck, Minefield or Fortification 
counters may not be overrun. 

H- 
The German SPA units (Maultier, Wespe and 
Hummel) may NOT make overrun attacks. 
Halftracks may not overrun Armored Vehicles 
(including enemy halftracks). 

I- 
Units may only be overrun when they are in 
clear terrain or clear terrain-road hexes. 

Close Assault Tactics 

All types of Russian and German infantry and 
engineer units as well as Russian cavalry units have 
the option of using Close Assault Tactics instead of 
making a normal attack. Close Assault takes place 
AFTER movement. 

A- 
Close-assaulting units must be directly adjacent 
to the defending unit or stack of units (i.e. in one 
of the six surrounding hexes). 

B- 
CAT attacks take place after all movement, 
normal attacking and overrun attacks are 
finished. 

C- 
Units utilizing CAT may NOT make normal attacks 
in the same turn. They may, however, move in the 
same turn. (NOTE: Overrun and CAT attacks are 
the only exceptions to the general rule which 
forbids movement and combat by the same unit in 
the same turn.) 

D- 
Close Assaulting units have their effectiveness 
increased by subtracting 2 from their die-roll result: 

e.g. a die-roll of “2” becomes a die-roll of “0”. The 
defending stack must be treated as one combined 
defense factor, and may suffer only one Close 
Assault per turn. 
E- 
If infantry and engineer units are stacked together 
when Close Assaulting the same defender, the 
effectiveness of that Close Assault is further 
increased by raising the odds in their favor to the 
next highest ratio (as in the Overrun rule). At least 
one engineer unit must be stacked with at least 
one infantry or cavalry unit, on at least one of the 
hexes of Assaulting units. 

F- 
Units capable of using Close Assault do not HAVE 
to use it to attack adjacent enemy units (they may 
attack them normally in the normal attack phase if 
the player so desires). 

G- 
Close Assault is the only way in which “I” units 
may attack Armored Vehicle units. 

H- 
Any type of defending unit may be attacked using 
Close Assault Tactics, CAT may be used in any 
type of terrain. 

I- 
Russian cavalry may not move more than one hex 
in a turn in which it is to be used for a Close 
Assault. No unit may use the road movement rate 
and make a Close Assault in the same turn. 

J- 
Halftracks may not use CAT. 

Weapon-to-Target Relationships 

The class-key letter symbols determine what type of 
weapons a unit is armed with. 

“I” = Infantry weapons (rifles \& machineguns) Note: 
Halftracks have “I” type weapons. 

“A” Armor Piercing Weapons (high velocity tank \& antitank 
guns). 

“H” = High Explosive Shells (low velocity howitzers, 
etc.). 

“M” = Mortars (similar to “H”). 

Page 6 of 12 


WEC Weapons Effectiveness Chart 

The effectiveness of these weapons changes in 
relation to target type and range from target. This is 
reflected in the WEC by doubling and halving a 
unit’s attack factor accordingly. Note: Half range is 
always rounded off to the nearest whole hex; the 
attacking unit loses fractional RF’s (for example: half 
of 9 is 4). 

A- 
When making a normal Combination Attack 
against a mixed stack of units, determine what 
type of target (Armored or Non-armored) 
predominates in the stack and then treat the 
entire stack as if all the units in it were that type 
of target. If the target stack is divided evenly 
between Armored and Non-armored targets, 
treat the whole stack as if it were that type of 
target least favorable to the particular attacking 
unit(s). 

B-Units in towns are considered as Armored 
targets whether or not such units actually are 
armored. Units stacked together in towns MUST 
be attacked as one combined DF. 

C- 
Fortifications are considered as Armored 
targets. 

D- 
WEC is not used in determining overrun odds. 

Obstacles and Elevations 

The PanzerBlitz mapboard is a two-dimensional 
representation of a three dimensional space. The 
various terrain features, aside from affecting 
movement and defense, also affect the ability of 
attacking units to fire at given defending units. Since 
the weapons used in PanzerBlitz are direct-fire 
weapons, an attacking unit may not fire at any target 
which it cannot “see.” The terrain symbols on the 
map show the location of potential obstructions and 
the obstacle-hex side symbols show in which 
direction fire is obstructed. These “heavy” hex-side 
symbols are color coded according to the type of 
obstruction they represent. 

There are three general types of obstacle/hex-sides 
which cut off the line of sight and therefore prevent 
fire: 

1. LOW OBSTACLES: Ground-level Green (woods) hex-
sides and ground-level Gray (town ) hex sides. 10 to 
20 meters. 
2. MEDIUM OBSTACLES: Dark Brown (Slope) hex-sides. 
60 to 70 meters. 
3. HIGH OBSTACLES: Orange (hilltop hex-sides. 100 to 
140 meters. 
Whether or not a firing unit can “see over” these 
Obstacles depends upon the elevation at which the 
firing unit and its potential target are. Units themselves 
are not considered as obstacles, and players may fire 
over or through all units, enemy or friendly. 

There are three elevations at which a unit may be: 

1. GROUND LEVEL: 0 meters. 
2. SLOPE LEVEL: 50-60 meters. 
3. HILLTOP LEVEL: 100 to 140 meters. 
The elevation at which a unit is, is of course 
determined by the terrain it is on. The Target Elevation 
Table shows, in a general way, what hex-side symbols 
obstruct the line-of-fire in different cases. The table, 
however, does not cover all the situations which might 
arise. 

How to Determine the Line of Sight/ Line 
of Fire (LOS/LOF) 

For all practical purposes the Line-of-Sight is 
equivalent to the Line-of-Fire. In “real life” the Line-of-
Sight would be a perfectly straight line. In the game, 
however, the Line-of-Sight is traced through the 
hexagons in a manner similar to the way units are 
moved. First, determine the range to the target. Then 
plot the route through the hexagons which your Lineof-
Sight takes. The Line-of-Sight should trace through 
the exact same number of hexes as the range to the 
target. The Line-of-Sight, in other words, should be as 
short and as straight as possible while still conforming 
to the hexagonal grid. Whenever there is more than 
one possible “shortest and straightest” route, choose 
the one least favorable to the attacker: i.e. if one route 
is clear and an alternate route is obstructed, the 
obstructed route is considered to be the one through 
which the Line-of-Sight must be traced. In other words, 
the defender gets the benefit of the doubt. Using the 
TET and the special notes, determine whether or not 
any hex-side symbols encountered in tracing the Lineof-
Sight, actually obstruct the Line-of-Sight. 

Page 7 of 12 


Special Notes (To Be Used In 
Conjunction With TET) 

A- 
When firing FROM a slope or hilltop TO a 
ground-level target, the Line-of-Sight is 
obstructed if the target unit is directly behind a 
Gray or Green hex-side: i.e. “directly behind” 
means that the intervening Gray or Green hex-
side(s) forms one or more of the hexagon sides 
of the target hex itself. 

B- 
When firing FROM ground-level TO a target on 
a hilltop or a slope, the Line-of-Sight is 
obstructed if the FIRING UNIT is directly behind 
a Gray or a Green hex-side. 

C- 
When firing FROM a hilltop TO a ground-level 
target, the Line-of-Sight is obstructed by 
intervening Brown hex-side symbols if such a 
symbol is closer to the TARGET UNIT than to 
the firing unit or if the symbol is exactly mid-way 
between the two. To determine the relative 
position of the Brown symbol, count the number 
of hexagon SIDES through which the Line-of-
Sight is traced (including the side of the firing 
hex and the side of the target hex). 

D- 
When firing FROM ground-level TO a target on 
a hilltop, the Line-of-Sight is obstructed by 
intervening Brown hex-side symbols if such a 
symbol is closer to the FIRING UNIT than to the 
target, or exactly midway between the two. 
(Note B is the converse of Note A and Note D is 
the converse of Note C.) 

E- 
The ONLY case in which a unit may trace an 
unobstructed Line-of-Sight through more than 
ONE Orange (hilltop) hex-side symbol is when 
both the target and the firing unit are on hilltops. 
In all other situations, the LOS is obstructed if it 
must be traced through more than one Orange 
hex-side. 

F- 
No matter what the obstacle or the terrain, a unit 
may ALWAYS fire at a target to which it is 
directly adjacent (regardless of elevation). 

G- 
Notice that in some cases there are towns and 
woods on top of hilltops. The Green and Gray 
symbols in these cases obstruct ALL fire, no 
matter what the elevation of the target and the 
firing unit (except when directly adjacent to each 
other as per note “F”). 

H- 
Note that when both the target and the attacker 
are on ground-level ALL hex-side symbols 

(Gray, Green, Brown, and Orange) obstruct the 
Line-of-Sight (except as per note “F”). 

I- 
(MAPBOARD NOTES:) The interior hexes on the 
large plateau-like hilltop on board No.2 are all 
hilltop hexes even though they do not have 
Orange hex-sides superimposed upon them. For 
practical and esthetic purposes the board designer 
felt it would have been redundant to so outline 
those hexes. You will notice that in some cases, 
slope hexes do not have Brown hex-side symbols. 
This is because the Brown symbols actually 
represent the way in which slopes curve and form 
“corners” to obstruct the LOS. Consequently, 
whenever a slope is relatively straight, it does not 
obstruct the LOS along it. In some instances, two 
or more slope hexes meet without a hilltop being 
formed between them. This represents a ridge or 
“razorback” hillock. The Brown hex-side symbols 
represent the “spine” of these ridges. 

Hill and Slope Defense Exceptions 

As indicated on the TEC (Terrain Effects Chart) a unit 
attacking an enemy unit defending on a slope or 
hilltop, attacks at half-attack-factor. There are some 
exceptional cases, however, in which the attacker is 
NOT halved. They are: 

When the defending unit is on a hilltop, an attacking 
unit is NOT halved if it is also on a hilltop (not 
necessarily the same hilltop). A defender on a hilltop 
could conceivably be attacked by units not on hilltops 
and units on hilltops as part of the same attack; in 
which case the attackers not on hilltops would STILL 
be halved. 

A- 
When a defender is on a slope, the attacking unit 
is NOT halved if the attacker is directly adjacent to 
the defender (regardless of elevation). If, however, 
there is a Brown hex-side symbol between an 
adjacent attacker and defender, the attacking 
unit’s factor IS halved. Units on slopes can 
conceivably be attacked by attacking units which 
are halved and which are not halved as part of the 
same attack. 

B- 
In all situations other than those described in “A” 
and “B” above, an attacking unit (regardless of 
elevation is halved when firing at units defending 
on slopes or hilltops. 

Page 8 of 12 


Gullies and Streambeds 

The gullies and streambeds shown on the board are 
DEPRESSIONS (minus 5 to 7 meters). The terms 
“streambeds” and “gullies” are interchangeable. 

A- 
Units in gullies may not fire at (or BE fired at by) 
units at ground level or in other gully-hexes 
(unless they are directly adjacent to each other). 

B- 
Units in gullies may fire at (and BE fired at by) 
units on slopes and hilltops. In these cases treat 
the unit in the gu7lly as if it were at ground-level 
for TET purposes. 

C- 
“Fords” represent exposed (not depressed) 
areas of a streambed. They are equivalent to 
clear terrain for all purposes. The hexagons on 
which a road crosses a gully are also 
considered as clear terrain (even if for some 
reason the road becomes unusable). 

D- 
Hexagons containing the End or beginning of a 
gully are treated as full fledged gullies. 

Spotting 

When a defender is in a Woods or a Town hex, he 
may not be fired upon by units which are not directly 
adjacent to him unless he has been “spotted.” 

A- 
To “spot” a unit in a town or woods hex, the 
attacker must have a friendly unit directly 
adjacent to the defender. 

B- 
The unit doing the spotting does not necessarily 
have to be involved in the attack upon the 
spotted unit. 

C- 
Dispersed units may not be used for spotting. 

D- 
The same unit may spot any number of adjacent 
enemy units. 

Wreckage 

Whenever an ARMORED vehicle unit (including 
halftracks) is destroyed in combat, remove it from 
the board and replace it with a Wreck counter. 

A- 
Wrecks may not be moved or removed. 

B- 
Wrecks count as a unit for stacking purposes. 

C- 
A Wreck on a road NEGATES THE ROAD, and 
that hex is treated as a non-road hex. Vehicle 
units may not enter swamp/road hexes 

containing Wrecks, nor may they cross Green-hex 
sides from a road hex with a wreck in it. 

D- 
The presence of Wrecks has no effect on combat. 

E- 
It is possible to have as many as three German or 
two Russian Wrecks in a hex, or a combination of 
German and Russian Wrecks not exceeding a 
total of THREE. 

F- 
Units may enter hexes containing enemy and/or 
friendly Wrecks as long as they do not exceed 
stacking limits. 

Mines 

A- 
The player with the minefield pieces positions 
them anywhere on the board he desires unless 
otherwise directed by the Situation Card. One 
mine per hex. 

B- 
Once positioned, mines may not be moved. 

C- 
Minefields have no friends – they affect both sides. 

D- 
As soon as a unit moves onto a minefield, it must 
stop. 

E- 
The opposing player, during the combat portion of 
his turn rolls the die for the “attacking minefield. 
The minefield attacks ALL units at 2 to 1 odds, no 
matter what the terrain. Surviving units may move 
off in there next turn. 

F- 
A minefield is never “used up.” It remains active 
until removed by an Engineer unit. 

G- 
Engineers remove mines by moving adjacent to 
them and on the turn after moving adjacent rolling 
the die. A roll of “1” removes the mines. If 
unsuccessful, the Engineers may remain adjacent 
roll each turn until they get a “1”. 

H- 
Units “dispersed” by minefields may not move off 
them, and suffer minefield attack in their next turn 
again. 

I- 
Minefields do not count against staking limits. 

J- 
Minefield “attacks” take place BEFORE normal 
attacks. 

Page 9 of 12 


Positional Defenses 

Blockage 

The BLOCK counters represent tank traps, road 
blocks, barbed wire, felled trees and anything else 
that might impede movement. 

A- 
Blocks may be placed anywhere on the board, 
no more than one Block per hex. Once placed 
they may not be moved or removed. 

B- 
A unit may only enter a hex with a Block in it if it 
begins its turn directly adjacent to the Block. 
Upon entering the Block-hex, the unit must stop 
and go no further that turn. In their NEXT turn 
they may move off the Block at the normal 
movement rate. 

C- 
Blocks do not obstruct the Line-of-Sight and 
have no effect upon combat. 

D- 
When a Fortification is destroyed, replace it with 
the Block counter. 

E- 
A Block counter on a road negates the road in 
that hex. VEHICULAR UNITS may not enter 
Block-road hexes in Swamps nor traverse 
directly adjacent Green hex-sides when moving 
off a block road hex. 

Fortifications 

(Bunkers, Redoubts, prepared trenches, etc.) 

A- 
Fortifications are place wherever on the board 
the player desires (except swamp hexes) or as 
directed by the Situation Card. Once placed they 
may not be moved. 

B- 
Fortifications do not affect movement or 
stacking. 

C- 
Units in fortifications defend using the defense 
factor of the fortifications only. Their own DF’s 
do not count. Fortifications are treated as an 
ARMORED target. Terrain and Weapons 
Effectiveness are taken into account when a 
fortification is defending. Any type of unit(s) may 
occupy a fortification within normal stacking 
limitations. 

D- 
Units may fire from fortifications using their 
normal attack factor (AF). 

E- 
If a fortification is destroyed, any units in it are 
also destroyed. 

F- 
The fortification counter itself, has no attack factor 
and it my only defend. 

G- 
If abandoned or unoccupied, fortifications may be 
“captured” and used by the opposing player. To 
capture a fortification, simply move a unit into the 
unoccupied fortification counter. 

H- Unwanted fortifications may only be destroyed by 
attacking them with one’s own fire weapons. 

I- 
The fortification unit is not an obstacle to fire, 
whether units are in it or not. 

J- 
Units are indicated as being IN the fortification by 
placing them UNDER the fortification counter. 
Friendly units ON TOP of fortifications (i.e. 
“outside” the fortification) and the fortification 
counter on which they are sitting, may be attacked 
using any one of the three Normal attack 
techniques as well as Close Assault Tactics. 

K- 
Enemy units may be actually ON the fortification 
counter while friendly units are still in it. In this 
case you may fire on that hex but you must roll the 
die twice for this attack, once for the attack on the 
enemy units and then for the effect on your own 
fortified units. In this case it is possible to destroy 
the enemy unit and not your own, or both. Or you 
may simply have a combination of dispersals. 
Enemy units may not enter a fortification hex if 
there are friendly units ON TOP of the fortification. 

L- 
When enemy units are sitting on top of a 
fortification, friendly units inside the fortification 
may not leave. The friendly units may still attack 
any enemy units within range, including the units 
sitting on the fortification: when attacking the units 
sitting on the fortification the units inside attack as 
if they were adjacent to the enemy. Similarly, the 
enemy units may attack the fortification as if they 
were adjacent to it. The units inside, in this case, 
may NOT use Close Assault Tactics; the enemy 
units on top may use CAT. 

M- Fortifications may suffer “dispersal” in which case 
the units occupying them are dispersed also. 

N- 
No more than one fortification may be placed in a 
given hex. 

O- 
Armored vehicles may NOT make Overrun Attacks 
against fortifications. 

P- Destroyed fortifications are replaced with block 
counters. 

Q- 
Fortifications may NOT be placed on top of Mines. 

Page 10 of 12 


\section{Game Procedure}

\subsection{Sequence of Play}


The game is played in turns, each player moving 
and having combat sequentially. Two “Player-Turns” 
equals one complete “Game-Turn”. 

\subsection{German Player Turn}

Step 1 German player resolves any Minefield 
attacks against Russian units. 

Step 2 German player announces which of his units 
are attacking which Russian units, and what 
attack techniques are being used. 

Step 3 German player resolves all Normal combat, 
rolls the die once for each attack. German 
player flips face-down all firing units, as they 
are fired, to signify that they may not move. 

Step 4 German player moves as many face up 
VEHICULAR units as he desires, executing 
any Overrun attacks as he does so. 

Step 5 German player moves any face-up NONVEHICULAR 
units and makes Close Assaults 
after doing so. 

Step 6 German player turns ALL his units FACEUP, 
including those dispersed by Russian 
attacks in the previous turn. 

\subsection{Russian Player Turn}

Step 7 The Russian player repeats Steps 1 through
6 using his own units.

Step 8 Indicate the passage of one complete
Game-Turn on the Turn Record.

Players repeat steps 1 through 8 for as many turns
as the Situation Card indicates or until one player
concedes.


\section{Optional Rules}

Players may employ as many or as few of the 
Optional Rules as desired. 

Indirect Fire 

German SPA units (Maultier, Wespe and Hummel) 
and all German and Russian mortar units 9M) may 
use Indirect Fire: i.e. they may fire at units which they 
cannot “see”; firing over all obstacles to the limit of 
their ranges. 

A- 
Indirect Fire may only be used in conjunction with 
CP units (200’s). 

B- 
The CP unit must be able to see the target: i.e. 
trace a clear Line-of-Sight to the target. 

C- 
Targets in towns or woods must still be “spotted” 
(but not necessarily by the CP unit itself). The 
spotting unit would radio or flare-signal the CP, 
which in turn would radio the target information to 
the unit firing indirectly. 

D- 
SPA units may only use Indirect Fire against 

targets which are at more than half the SPA units’ 

range. 

Real Space Line of Sight Determination 

Use a straight-edge (ruler, stiff cardboard, etc.) to 
determine the Line of Sight. Determine range in the 
normal way but to calculate Line of Sight, place a 
straight edge between the center of the firing hex and 
the center of the target hex. Only those hex-side 
symbols intersected by the straight-edge need be 
taken into account. If the straight edge bisects a hex-
side symbol through its LENGTH, take that symbol 
into account UNLESS it is a BROWN symbol which 
connects with an Orange symbol. The defender gets 
the benefit of the doubt if the straight-edge cuts 
exactly through the “corner” of a hex where a symbol-
side and a non-symbol side meet. In any other 
ambiguous cases, use “common-sense” to decide, 
keeping in mind the fact that the board is representing 
a three-dimensional space. 

Panzerblitz Assault 

Infantry units riding on Armored vehicles may “jump 
off” in the hex immediately in front of a unit about to be 
Overrun by those Armored vehicles. The infantry may 
then Close Assault the units which were just Overrun 
by the vehicles. 

Page 11 of 12 


Ammunition Rule 

Anytime he desires, a player may fire one or more of 
his “H” or “M” class units “intensively.” Intensive fire 
allows such units to TRIPLE their normal attack 
factor. 

A- 
A given unit may fire intensively only once per 
game, immediately after which it is removed 
from play (its ammunition has been expended 
and its gun tubes burned out). 

B- 
Units removed from play under this rule are 
counted as units lost. 

C- 
Players may exercise this intensive fire option as 
many times as they wish during a game, limited 
only by the quantity of “H” and “M” units 
available. 

D- 
“H” CLASS ARMORED VEHICLE UNITS MAY 
NOT USE INTENSIVE FIRE AS PART OF AN 
OVERRUN ATTACK. 

E- 
MAULTIER unit may not use intensive fire. 

Experimental Rules 

The following experimental rules are just that: 
experiments! The game factors and mapboard are 
not necessarily designed to accommodate them. 

Hidden Deployment 

Utilizing one of the low unit-count Situations or one 
of your own devising, allow one player to secretly 
deploy his units on the board, marking their positions 
on separate pieces of paper using the Map Location 
System, and remove the units from the board. The 
other player is not shown where a given enemy unit 
is until it fires its weapons or until the second player 
has a clear Line of Sight to the unit’s position. Once 
hidden units are moved, fired or spotted, they must 
be placed on the board and left in view. Minefields 
may be hidden anywhere; combat units may only be 
hidden in woods, swamps, towns, or behind hills. 

Experimental Indirect Fire 

Allow any German unit(s) to perform the fire-
direction function of a CP unit. Allow Russian 
Guards infantry this capability. 

Impulse Movement \& Return Fire 

This rule is a “meshing” of the two player turns which 
breaks down the simulated event into smaller bits of 
time called “impulses.” Players use the following 
sequence for a complete turn: 

gA German 
attacks. 
moves units, executes Minefield 
gB Russian fires (do not flip firing units). 
gC German fires (does flip firing units). 
gD German moves non-firing units, executes 

Overruns. 

gE 
German makes Close Assaults, flips ALL units 
right-side-up. 

rA 
Russian moves units, executes Minefield 
attacks. 

rB 
German fires (do not flip firing units). 

rC 
Russian fires (does flip firing units). 

rD 
Russian moves non-firing units, executes 
Overruns. 

rE 
Russian makes Close Assaults, flips ALL units 
right-side-up. 

In this system, all units with a movement factor of 
MORE than 1 may only move HALF their movement 
factor in any movement impulse. Units with an ODD 
movement factor use the greater “half” of their factor in 
impulse “D.” Units with a movement factor of 1 may 
move one hex EACH movement impulse (A\&D) but 
they may not move at the road movement rate in 
impulse “A.” 


\section{House rules}

\subsection{Artillery-supported CAT}

``75mm infantry guns (2 H 12) and company mortars (3 M 12) may add their factors
in to CAT infantry attacks if within range. Infantry guns require LOS; mortars
may fire indirectly.'' 
\href{http://boardgamegeek.com/article/15828184#15828184}{CAT support}




\end{document}
