
From href{http://http://grognard.com/errata/st/s&t81.txt}{Tito errata at Grognards}

Tito addenda

[7.22C] (change) Defecting Chetniks are placed in the
Yugoslav/partisan Mountain triangle of the occupied zone.

[7.23B] (change) +1 if Italian Surrender (10.3, not Withdrawal)
has taken place.

[7.42] (addition) Within the limits of the Sequence of Play
(4.0) and 7.68 (limitations on the number of counters), the
Yugoslav player is free to choose the order in which his
reinforcements may be created.

[7.44E] (addition) No Chetnik unit may generate reinforcements or
Victory Points if an Objective Display also is occupied by
Pro-Axis Chetniks .

[7.49] (clarification) The halvings apply to such factors as
weather, etc., and not to a Recruitment Value of 1/2 of a
display.

[7.63A, 7.67A] (changes) No larger guerrilla unit may be created
in a Hideaway circle.

[8.24B] (change) When engaged in combat entirely against
guerrilla units in a Mountain Triangle, Axis units incurring a
retreat result are not obligated to retreat.

[8.55] (change) The intrinsic Defense Strength of the four cities
are not automatically eliminated by Italian Surrender. (They
reflect historical Italian units that remained loyal to the Axis
after the surrender.)

[8.6] (change to CRT) Modifications for Tito, Anti-Guerrilla
Operations, and mountain units are made to the probability ratio
(8.21B) and are not column shifts on the Combat Results Table.
Thus, after all such modifications have been made, attacks at
odds lower than I to 2 are not permitted, and attacks at greater
than 12 to 1 are resolved on the 12 to 1 column.

[9.43] (addition) If the 501st is participating in an
Anti-Guerrilla Operation, it need not be removed from the map
during the Planning Segment (8.42), although the Axis Player
secretly makes a notation as to which operation the 501st is
committed. The unit is deployed normally according to 8.22.

[10.22] (change) Delete the reference to Chetnik Collaboration
die rolls.

[10.21] (change) Italian units may retreat into the Croatia
Mountain Triangle, but must move out in the ensuing Axis Movement
Phase.

[10.3] (addition) Chetnik Collaboration die rolls (7.2) are
affected by Italian Surrender.

[11.22A] (addition) Normal weather prevails on Game-Turn 1.

[13.62] (change) The Yugoslav player is eligible for
Pro-Yugoslav Chetnik replacements between Game-Turns 2 and 8 (not
13).

[13.92, 14.3E] (clarifcation) The Yugoslav player loses 5 Victory
Points for each optional Soviet reinforcement unit used on
Game-Turn 15.

[14.3C] (change) The exception to this case should read: "After
Allied Support Withdrawal (not Game-Turn 13), this Case should be
ignored."

[14.3E] (change) The Yugoslav player loses 20 Victory Points (not 10)
for each Soviet and pro-Yugoslav Bulgarian unit on the map during the
Victory Point Stages of Game-Turns 16 and 17.

[14.4] (addition) Add Case 9.41 to those listed.

[15.5] (change) Add l00 points to each level of victory For example,
a Yugoslav Decisive Victory requires greater than 800 (not 700) Points,
and an Axis Decisive Victory requires less than 401 (not 301)
Points.

[15-6] SHORTENED GAME (Optional)
The game ends at the completion of the Victory Point Stage of
Game-Turn 15. Victory is determined according to the original schedule
contained in Case 15.5 of the Rules booklet

COUNTER ABBREVIATIONS

Croat
Ust: Ustacha (political militia); Gd: Guard.

Italian
CD Alpi: Cacciatori delli Alpi; Alpi Gr: Alpi Graie; Lombard: Lombardia.

German
SS: Schutzstaffel (elite troops); LW: Luftwaffe (air force field troops)
L: Leight (light); Cos: Cossack.

